Forecasting the black Sigatoka development rate:\\ A comparison of machine learning techniques

### 2017-04-29 intro.tex
28 abril

This biological warning system is based on weekly observations of the
disease progression on young leaves of the plant.
%
\figref{fig:diseasestages} shows an example of three progressive
stages of the black Sigatoka.

\begin{figure}[ht] 
\centering
\begin{tabular}{c@{\;}c@{\;}c}
  \includegraphics[width=.32\linewidth]{Roya_a} &
  \includegraphics[width=.32\linewidth]{Roya_b} &
  \includegraphics[width=.32\linewidth]{Roya_c} \\
  (a) & (b) & (c) 
\end{tabular}
\caption{Examples of three disease stages of the black Sigatoka. 
(a)~Initial stage. (b)~Intermediate stage, and (c)~Advanced stage.} 
\label{fig:diseasestages} 
\end{figure}

@@

In this work, we compare five machine learning techniques to predict
the development rate of the black Sigatoka disease: support vector
regression (SVR), echo state networks (ESN), ridge regression,
elastic-net regression and ordinary least squares linear regression,
using input variables such as \review{completar las variables de
  entrada usada} to predict \review{completar las variables de salida}.

\todofn{The main contribution of this work}%
{Definitivamente ese no es el aporte principal del artículo.  Cuando
  esté listo el resto hay que volver aquí.  Creo que va a ser algo
  como mostrar la relevancia de la etapa de preprocesamiento, o
  mostrar que a pesar de ser sistemas complejos, con variables
  caóticas, lo mejor en este caso son predictores lineales, o algo por
  el estilo.} %
% 
 is the selection of the best machine learning technique
to forecast the black Sigatoka development rate according to the
metric indicated below.

The outline of the paper is as follows: Section~\ref{sec:related}
presents related works and Section~\ref{sec:techs} summarizes the
machine learning techniques selected for the analysis. In
Section~\ref{sec:data} we present the methodology used in this study
and describe data used for its verification.  The results and their
discussion are presented in section~\ref{sec:results}.  The
Section~\ref{sec:concl} concludes this article and presents lines for
future works.

>>>>

This biological warning system is based on weekly observations of the
disease progression on young leaves of the plant.
%
\figref{fig:diseasestages} shows an example of three progressive
stages of the black Sigatoka.
%	 
\begin{figure}[ht] 
\centering
\begin{tabular}{c@{\;}c@{\;}c}
  \includegraphics[width=.32\linewidth]{Roya_a} &
  \includegraphics[width=.32\linewidth]{Roya_b} &
  \includegraphics[width=.32\linewidth]{Roya_c} \\
  (a) & (b) & (c) 
\end{tabular}
\caption{Examples of three disease stages of the black Sigatoka. 
(a)~Initial stage. (b)~Intermediate stage, and (c)~Advanced stage.} 
\label{fig:diseasestages} 
\end{figure}

@@

In this work, we compare four machine learning techniques to predict
the development rate of the black Sigatoka disease: support vector
regression (SVR), echo state networks (ESN), elastic net regression 
and ordinary least squares linear regression,
using input variables: maximal air temperature, minimal air temperature,
 mean air temperature, mean relative humidity, minimal relative humidity, 
 maximal relative humidity, mean solar radiation, sum of precipitation, 
 maximal wind speed and mean speed wind, to predict the evolution stage 
 in the biological warning system.

 The main findings were: 1) The highest R2 was 60\% 2) The highest R2 
 were reached with linear models like support vector regression with 
 linear kernel, 3) As little as three meteorological variables can 
 be used because of the correlations detected among variables.

The outline of the paper is as follows: Section~\ref{sec:related}
presents related works and Section~\ref{sec:techs} summarizes the
machine learning techniques selected for the analysis. In
Section~\ref{sec:data} we present the methodology used in this study
and describe data used for its verification.  The results and their
discussion are presented in section~\ref{sec:results}.  The
Section~\ref{sec:concl} concludes this article and presents lines for
future works.
%

### 2017-04-29 methodology.tex
2016-12-09-01
28 abril

The meteorological stations of Corbana acquire data every five
minutes.
%
\todofn{However, for the current study weekly data were used.}
{Hay que explicar por qué? $\rightarrow$ reducción de ruido, etc.}
%
computed on the data collected by nearby stations in each farm. Experiments were carry out with daily periodicity in meteorological variables and the results proved do not improve the prediction. Besides, weekly data pretend to diminish noise due sensor accuracy, missing values and outliers no detected.

@@

In the farm 28 Miles 1\% and in La Rita 2.25\% of the data were missing.
%
To fill-in the missing values \todofn{spline interpolation}{falta una
  referencia} was used.
%
The data collected did not exhibit outliers.

Due the fact that the variables measure meteorological or biological
processes, they are discretized \todofn{in order to reflect data trends}%
%
{Formalmente la discretización NO captura tendencias, sino la derivada
  de la función en el tiempo.  Para que no se capture ruido uno
  simplemente usa un filtro pasabajas primero (los estadísticos le
  dicen diferente, pero no sé cómo).  En otras palabras, si querés la
  tendencia, con usar la derivada de un filtro gaussiano sobre los
  datos ya se ve la tendencia muy bien con el signo de esa derivada.
  Con la varianza del filtro uno ajusta qué nivel de detalle le
  interesa, o lo qué es lo mismo, cuanto tiempo quiere usar para el
  pronóstico de tendencias (variable crece, está igual o decrece).  El
  valor de salida de la derivada filtrada también se puede discretizar
  si solo interesa esa tendencia de forma difusa, pero quizá para la
  regresión es mejor mantenerla continua (si está filtrada, con un
  filtro o mejor aún con un banco de filtros, al estilo wavelets).

  Eso es TOTALMENTE independiente de la cuantificación.  Ahora, un
  cuantificador tiene en principio un efecto de filtro pasabajas, pero
  eso reduce el ruido y en realidad no dice nada de tendencias.  En
  resumen, formalmente la cuantificación en sí no refleja nada de un
  comportamiento temporal, y ``tendencia'' es en principio un
  comportamiento temporal!}.
%
The variation range of each variable is uniformly discretized.
%
This discretization removes the effect of small variations in the data,
either by inaccuracies of the instruments (meteorological variables)
or by subjective bias introduced by the human collecting the data
(biological warning system).

The coefficient of variation
$C_v(x)={\sqrt{E((x-E(x))^2)}}/{E(x)}$ of each variable $x$ is
used to determine the number $n$ of discretization levels as
\todofn{$n=\lfloor 100 \ C_v(x) \rfloor$,}%
%
{Hay algo que no entiendo.  ¿De dónde sale el 100?  ¿Por qué 100?
  Así como está esto, el mismo rango de variación de temperatura
  produce diferentes $n$, con solo un cambio en el promedio!  Eso es
  muy extraño.  Por otro lado, los rangos de las variables son tan
  diferentes, que el número $n$ va a terminar siendo casi cualquier
  cosa.  Yo hubiese esperado que la desviación estándar se use para
  discretizar las variables.  Otro problema formal más serio, es que
  el coeficiente de variación solo se puede usar/tiene sentido si la
  escala de medición es de razón, y la temperatura, la precipitación y
  la humedad son usualmente mediciones de intervalo! Los $L_i$ ni tan
  siquiera sé qué son porque creo que son algo subjetivos a la persona
  que los toma ¿no?, aunque creo que iban con el porcentaje de área
  afectado, lo que también sería medición de intervalo!}
%
where $\lfloor \cdot \rfloor$ is the round operator.

Each feature \todofn{was scaled}{¿escaladas cómo? ¿$sx$? ¿$sx+b$?
  ¿$sf(x)$?} to fit in the range 0 and 1. The variable $E_s$ to be
predicted was not scaled.

\TODO{Aquí faltan la fórmula general de normalización usada.  Lo que
  sigue lo estoy asumiendo, pero puede ser que no sea así}

@@

\TODO{Explicar brevemente la razón de la normalización, ojalá con una
  referencia bibliográfica}

@@

Given $n$ records $y_i$, $i=1\ldots{}n$ of the actual outcome of a
process. The mean $\bar{y}$ of the observed data is given by
\begin{equation*}
  \bar{y} = \frac{1}{n} \sum_{i=1}^{n} y_i
\end{equation*}
Let $\hat{y}_i$ be the predicted value for $y_i$. Then, the mean
square error (MSE) $S_e^2$ and the \review{¿cómo se llama esto?
  ¿unexplained variance?} $S_R^2$ are estimated as
\begin{align*}
S_e^2 &= \frac{\sum_{i=1}^{n} {(y_i-\hat{y}_i)}^2 }{n} &
S_R^2 &= \frac{\sum_{i=1}^{n} {(\hat{y}_i-\bar{y})}^2 }{n}
\end{align*}
The root mean square error is defined as $RMSE = \sqrt{S_e^2}$ and the
coefficient of determination is
\begin{equation*}
R^2 = \frac{S_R^2}{S_R^2 + S_e^2}
\end{equation*}

\TODO{Favor revisar.  Vi una representación en términos de $R^2=1 -
  MSE/\sigma^2$ (con $\sigma^2$ la varianza), que no sé si está bien o
  no, pero lo mejor será poner la fuente de donde se tomó la anterior
  definición.  Había en el $S_R^2$ un $\bar{y}_i$ pero el promedio no
  depende de $i$ por lo que no tenía sentido.  Supuse que era el
  promedio, pero no sé.  El cuadrado en las $S^2$ ¿está bien?, o
  indirectamente estás diciendo que $RMSE=S_e$, que sería una mejor
  notación matemática (dejando RMSE solo como abreviatura textual)!}

@@

We use the Python programming language \todofn{with the Integrated
  Development Environment (IDE)}%
{El IDE es para la reproducción de resultados irrelevante, y por lo
  tanto información circunstancial que es mejor evitar.  Lo relevante
  es lo que tenga efectos directos en los resultados.  Si yo prefiero
  usar un editor de texto corriente para programar, el restulado final
  es el mismo, si uso el mismo compilador/intérprete.  Eso es la
  información relevante: ¿qué versión de Python concretamente? ¿Se
  compiló el Python o se usó interpretado?  ¿Cuáles versiones de
  Pandas y Numpy en concreto se usaron?} %
Spyder \citep{Continuum2015}, particularly with the libraries Pandas
\citep{mckinneypandas2010} and Numpy \citep{vanderWalt2011}.

The implementation for SVR, ridge and ordinary least squares
regressions in scikit-learn \citep{scikitlearn2011} were used.
%
Adjustments to the ESN implementation code of \citet{Lukose2012} were
necessary to allow its integration into our experimental
framework.

All experiments were performed on a PC computer with an
Intel\textsuperscript{\textregistered} Core i7-4800MQ processor,
2.70\,GHz, 16.0\,GB RAM, under the operating system MS Windows 8 Pro.

>>>>

The meteorological stations of Corbana acquire data every five
minutes.
%
computed on the data collected by nearby stations in each farm. Experiments were carry out with daily periodicity in meteorological variables and the results proved do not improve the prediction. Besides, weekly data pretend to diminish noise due sensor accuracy, missing values and outliers no detected.

@@

In the farm 28 Miles 1\% and in La Rita 2.25\% of the data were missing.
%
To fill-in the missing values spline interpolation was used. \TODO{Falta poner una cita de spline interpolation}
%
The data collected did not exhibit outliers.

@@

The variable $E_s$ to be predicted was not normalized.
This normalization step was made because learning schemes, like regression methods, 
deals only with ratio scales because they calculate the distance between two instances based on the 
values of their attributes \citet{Witten2011}.

@@

Given $n$ records $y_i$, $i=1\ldots{}n$ of the actual outcome of a
process. The mean $\bar{y}$ of the observed data is given by
\begin{equation*}
  \bar{y} = \frac{1}{n} \sum_{i=1}^{n} y_i
\end{equation*}
Let $\hat{y}_i$ be the predicted value for $y_i$. Then, the mean
square error (MSE) $S_e^2$ and the unexplained variance $S_R^2$ 
are estimated as \citet{Pennstate2017}
\begin{align*}
S_e^2 &= \frac{\sum_{i=1}^{n} {(y_i-\hat{y}_i)}^2 }{n} &
S_R^2 &= \frac{\sum_{i=1}^{n} {(\hat{y}_i-\bar{y})}^2 }{n}
\end{align*}
The root mean square error is defined as $RMSE = \sqrt{S_e^2}$ and the
coefficient of determination is
\begin{equation*}
R^2 = \frac{S_R^2}{S_R^2 + S_e^2}
\end{equation*}

@@

We use Python programming language in its interpreted 3.5.2 version,
 particularly with the libraries; Pandas (0.18.1 version)
\citep{mckinneypandas2010} and Numpy (1.11.1 version) \citep{vanderWalt2011}.

The implementation for SVR, elastic net and ordinary least squares
regressions in scikit-learn \citep{scikitlearn2011} were used.
%
Adjustments to the ESN implementation code of \citet{Lukose2012} were
necessary to allow its integration into our experimental
framework.

### 2017-04-29 related.tex
2017-01-13-prueba
28 abril

Several efforts have been made to apply machine learning techniques in
the automated discovery of relationships between environmental
variables and quantified descriptors for variables of agricultural 
interest such as the progress of diseases.
% 
\citet{Huang2010} summarize in their survey the
development of soft computing techniques in agricultural and
biological engineering, especially in the soil and water context for crop management and decision support in precision agriculture, including fuzzy logic, artificial neural
networks, genetic algorithms, Bayesian inference and decision trees. They do not present numeric results of each paper, only mention the main idea.
%
Similarly, \citet{Kim2014} survey more recent prediction
methods for crop pests using regression and machine learning approaches. Nor do they provide numerical results of each paper.
%
In general, the machine learning methods applied to predict the
evolution of plant diseases, can be classified in two main approaches:
1)~those whose main inputs are images, and 2)~Those whose main inputs
are environmental and biological variables. Our study focuses in the
second case.

% %
% For example, 
% 
% Me parece que aquí hay un error: Huang2010 es el survey, y sería mucha
% coincidencia que fuera lo mismo que el \citet{Glezakos2010} ¿o no?
%
% \cite{Huang2010} introduce a plant virus
% identification method based on neural networks with an evolutionary
% preprocessing stage, 
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Nota. la cita para lo de plant virus identificacion es  \citet{Glezakos2010}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\citet{Romero1995} relied on regression models using a stepwise
procedure to predict incubation and disease latency periods for the
black Sigatoka.
% 
He collected environmental data from two different farms in Costa Rica
between December 1993 and August 1995.
%
The prediction models reached coefficients of determination $R^2$ of
69\% or 78\% on the observed data for the incubation and disease
latency periods, respectively; however, the cross validation on
independent data sets failed.
%
In contrast, our proposal presents a model that can be generalizable to other farms that have data.

@@

\citet{Alves2011} used geoinformation techniques to
develop predictive models in the study of risk areas to soybean rust,
coffee leaf rust, and banana black Sigatoka, under consideration of
Brazil’s climatic characteristics and the distribution of soybean,
coffee and banana crops.
%
Temperature and rainfall data were acquired for the period from 1950
to 2000, and simulated data were generated for 2020, 2050 and 2080
using the SRES~A2 climate change scenarios.
%
Using principal components analysis, a single variable was generated
as a linear combination of 57 input variables, in order to determine
an index explaining 87\%, 88\% and 90\% of the data
%
\todofn{variability}{variabilidad de qué? - MODIFICADO: de los datos de cada uno }  
%
of soybean, coffee and
banana crops, respectively, in municipal districts across Brazil.
%
The climatic model was used to generate the zoning of the three plant
diseases, using temperature and leaf wetness as input.
%
%Redunda con lo que sigue:
%Areas favorable for the diseases were plotted against the main coffee,
%soybean and banana growing areas in Brazil.
%
This methodology enabled the visualization of the changes in areas
favorable for epidemics under possible future scenarios of climate
change.
%
How intermediate result, they characterized the monociclic process of the black Sigatoka using  nonlinear regression. Although they do not present the detailed results, it no seem that they wanted to predict the progression of the black Sigatoka in one, two or more periods ahead, how we do.

@@

Although these studies have their contribution, none proposed the kind of preprocessing that we present, nor pose how to predict more than one period ahead without trying to predict climate.

% Este no calza, porque no se ve qué es lo de machine learning que usa...
%
%\citet{Zhao2013} present an intelligent agricultural
%forecasting system based on wireless sensor networks.

\TODO{MODIFICADO - Importantísimo es explicar aquí brevemente por qué esos otros
  trabajos no se aplican al caso de la Sigatoka, o qué es lo que han
  hecho mal, que este trabajo sí hace bien}

>>>>

Several efforts have been made to apply machine learning techniques in
the automated discovery of relationships between environmental
variables and quantified descriptors for variables of agricultural 
interest such as the progress of diseases.
% 
\citet{Huang2010} summarize in their survey the
development of soft computing techniques in agricultural and
biological engineering, especially in the soil and water context for crop management and decision support in precision agriculture, including fuzzy logic, artificial neural
networks, genetic algorithms, Bayesian inference and decision trees. They do not present numeric results of each paper, only mention the main idea.
%
Similarly, \citet{Kim2014} survey more recent prediction
methods for crop pests using regression and machine learning approaches. Nor do they provide numerical results of each paper.
%
In general, the machine learning methods applied to predict the
evolution of plant diseases, can be classified in two main approaches: 
1)~those whose main inputs are images, and 2)~Those whose main inputs
are environmental and biological variables. Our study focuses in the
second case.
%
\citet{Romero1995} relied on regression models using a stepwise
procedure to predict incubation and disease latency periods for the
black Sigatoka.
% 
He collected environmental data from two different farms in Costa Rica
between December 1993 and August 1995.
%
The prediction models reached coefficients of determination $R^2$ of
69\% or 78\% on the observed data for the incubation and disease
latency periods, respectively; however, the cross validation on
independent data sets failed.
%
In contrast, our proposal presents a model that can be generalizable to other farms that have data.

@@

\citet{Alves2011} used geoinformation techniques to
develop predictive models in the study of risk areas to soybean rust,
coffee leaf rust, and banana black Sigatoka, under consideration of
Brazil’s climatic characteristics and the distribution of soybean,
coffee and banana crops.
%
Temperature and rainfall data were acquired for the period from 1950
to 2000, and simulated data were generated for 2020, 2050 and 2080
using the SRES~A2 climate change scenarios.
%
Using principal components analysis, a single variable was generated
as a linear combination of 57 input variables, in order to determine
an index explaining 87\%, 88\% and 90\% of the data variability 
of soybean, coffee and 
banana crops, respectively, in municipal districts across Brazil.
%
The climatic model was used to generate the zoning of the three plant
diseases, using temperature and leaf wetness as input.
%
This methodology enabled the visualization of the changes in areas
favorable for epidemics under possible future scenarios of climate
change.
%
How intermediate result, they characterized the monociclic process of the black Sigatoka using  nonlinear regression. Although they do not present the detailed results, it no seem that they wanted to predict the progression of the black Sigatoka in one, two or more periods ahead, how we do.

@@

Although these studies have their contribution, none proposed the kind of preprocessing that we present, nor pose how to predict more than one period ahead without trying to predict climate.
%

### 2017-04-29 techniques.tex
26-08-2016
28 abril

In the prediction of the development rate of the black Sigatoka, we
compare techniques such as least squares or ridge regression, commonly
encountered in the agricultural literature with machine learning
methods such as support vector regression, elastic regression and echo
state networks, where the parameter space of each technique is also
taken into account.

@@

\TODO{Alex, no sé hasta donde hiciste en las pruebas ese ``arreglo''
  de que los datos tengan todos una entrada igual a 1.  De no ser así,
  el método no puede encontrar el offset, y sería como forzar que el
  modelo tenga que pasar por cero...  a menos por supuesto que la
  constante se haya sacado de la ecuación al optimizar...}

\subsection{Ridge regression}

In contrast to the OLSR, for the ridge regression (RR)
\citet{Hoerl1988} proposed to add a term to penalize large weights
into the error function
\begin{equation*}
  E(\vct{w}) = 
  \left\| \mat{X}\vct{w} - \vct{y} \right\|_2^2 + 
  \alpha \left\|\vct{w}\right\|_2^2 
\end{equation*}
where the parameter $\alpha>0$ controls how strong is the shrinking of
the estimates towards zero.  This shrinkage introduces some bias but
helps to reduce the variance of the estimate.
%
The solution of the optimization problem (\ref{eq:problem}) in this
case is given by $\hat{\vct{w}}=(\mat{X}^T\mat{X} +
\lambda\mat{I})^{-1}\mat{X}^T \vct{y}$.

@@

Instead of $L_2$ regularization prior
($\alpha\left\|\vct{w}\right\|_2^2$) included in the ridge regression,
\citet{Tibshirani1996} used an $L_1$ term ($\lambda
\left\|\vct{w}\right\|_1$) for his lasso estimator, which permits to
select a subset of the available features by zeroing the weights of
the deselected features.
%
If the dimension $d$ of the data is larger than the number $n$ of data
samples, lasso will select a maximum of $d$ variables.

@@

\TODO{Algo que no entiendo es que las elastic net son en realidad una
  generalización del ridge, Lasso y OLSR.  Así que en los experimentos
  con las selección adecuada de parámetros este método debería
  comportarse al menos igual o mejor que esos otros regresores!  No sé
  hasta donde sea justificable usar este Y los otros, porque este ES
  los otros...}

\TODO{Alex, veo que usas mucho \citep{scikitlearn2011} para los
  métodos, pero ese es solo el paper de un tool y no los proponentes
  originales de los métodos.  Usualmente uno hace referencia a algún
  artículo, libro o tutorial donde ojalá los que propusieron el método
  son los que lo explican.  Ahí metí entonces en las referencias otros
  artículos por ese motivo.}

@@

\TODO{Estoy seguro que los lectores van a solicitar reducir esto y
  dejar solo la referencia.  Dejé solo lo que considero relevante,
  pero podés reducirlo más, si querés.}

From the perspective of Support Vector Regression (SVR) the regression
function is usually formulated as 
\begin{equation}
\label{eq:svrfunc}
  \tilde{y} = f(\vct{x}) = \dotp{\vct{w},\vct{x}} + b
\end{equation}
The weights are selected in a convex optimization problem
\citep{Smola2004}:
\begin{align*}
  \text{minimize}   & \frac{1}{2}\|\vct{w}\|^2 + C\sum_{i=1}^n(\xi_i+\xi_i^\ast \\
  \text{subject to} & 
  \begin{cases}
    y_i - \dotp{\vct{w},\vct{x}_i} - b &\leq \epsilon + \xi_i    \\
    \dotp{\vct{w},\vct{x}_i} + b -y_i  &\leq \epsilon + \xi_i^\ast\\
    \xi_i,\xi_i^\ast               &\geq 0
  \end{cases}
\end{align*}
where $\epsilon$ is the maximal allowed deviation of the targets
$\tilde{y}_i$ from the responses $y_i$, the slack variables $\xi_i$
and $\xi_i^\ast$ allow to cope with otherwise unfeasible constraints
for the optimization problem, and the constant $C>0$ controls the
trade-off between the flatness of $f$ and the tolerance to deviations
larger than $\epsilon$.

@@

\TODO{Habría que insertar aquí los kernels utilizados en los experimentos}

@@

For a given training input signal $u(n) \in \setR^{N_u}$ a desired target
output signal $y^{target}(n) \in \setR^{N_y}$ is known.
%
Here $n = 1,\ldots,T$ is the discrete time and \todofn{$T$ is the number of
data points in the training dataset.}{No puede ser!}

The training seeks to learn a model with output $y(n) \in
\setR^{N_y}$, where $y(n)$ matches $y^{target}(n)$ as close as
possible, by means of the minimization of an error measure
$E(y,y^{target})$ such that it also generalizes well to unseen data
\citep{Lukose2012}.

>>>>

In the prediction of the development rate of the black Sigatoka, we
compare techniques such as least squares regression and elastic-net regression, commonly
encountered in the agricultural literature with machine learning
methods such as support vector regression and echo
state networks, where the parameter space of each technique is also
taken into account.

@@

@@

Instead of $L_2$ regularization prior
($\alpha\left\|\vct{w}\right\|_2^2$) included in the ridge regression (RR),
\citet{Tibshirani1996} used an $L_1$ term ($\lambda
\left\|\vct{w}\right\|_1$) for his lasso estimator, which permits to
select a subset of the available features by zeroing the weights of
the deselected features.
%
If the dimension $d$ of the data is larger than the number $n$ of data
samples, lasso will select a maximum of $d$ variables.

@@

@@

From the perspective of Support Vector Regression (SVR) the regression
function is usually formulated as 
\begin{equation}
\label{eq:svrfunc}
  \tilde{y} = f(\vct{x}) = \dotp{\vct{w},\vct{x}} + b
\end{equation}
The weights are selected in a convex optimization problem
\citep{Smola2004}:
\begin{align*}
  \text{minimize} \quad  & \frac{1}{2}\|\vct{w}\|^2 + C\sum_{i=1}^n(\xi_i+\xi_i^\ast) \\
  \text{subject to} \quad & 
  \begin{cases}
    y_i - \dotp{\vct{w},\vct{x}_i} - b &\leq \epsilon + \xi_i    \\
    \dotp{\vct{w},\vct{x}_i} + b -y_i  &\leq \epsilon + \xi_i^\ast\\
    \xi_i,\xi_i^\ast               &\geq 0
  \end{cases}
\end{align*}
where $\epsilon$ is the maximal allowed deviation of the targets
$\tilde{y}_i$ from the responses $y_i$, the slack variables $\xi_i$
and $\xi_i^\ast$ allow to cope with otherwise unfeasible constraints
for the optimization problem, and the constant $C>0$ controls the
trade-off between the flatness of $f$ and the tolerance to deviations
larger than $\epsilon$.

@@

Kernels used in this work were:

Linear kernel: 
\begin{equation*}
K(\vct{x}_i,\vct{x}_j)={\vct{x}_i}^T \vct{x}_j
\end{equation*}

Radial basis function (RBF): 
\begin{equation*}
K(\vct{x}_i,\vct{x}_j) = exp\left( \frac{ \| \vct{x}_i - \vct{x}_j \|^2}{2 \sigma^2} \right)
\end{equation*}
where $\sigma$ is the parameter of gausian model.

Sigmoid:
\begin{equation*}
K(\vct{x}_i,\vct{x}_j) = tanh \left[ -c + \frac{ \vct{x}_i  \vct{x}_j }{ \sigma^2} \right]
\end{equation*}
with $c \geq 0$  and $\sigma^2$ is the scaling vector.

@@

For a given training input signal $u(n) \in \setR^{N_u}$ a desired target
output signal $y^{target}(n) \in \setR^{N_y}$ is known.
%
Here $n = 1,\ldots,T$ is the discrete time and $T$ is the number of
data points in the training dataset.

The training seeks to learn a model with output $y(n) \in
\setR^{N_y}$, where $y(n)$ matches $y^{target}(n)$ as close as
possible, by means of the minimization of an error measure
$E(y,y^{target})$ such that it also generalizes well to unseen data
\citep{Lukose2012}.
%
